\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=0.8in]{geometry}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{listings} % linebreaks in lstlisting
\usepackage{minted} % python code


\lstset{
basicstyle=\small\ttfamily,
columns=flexible,
breaklines=true
}
\begin{document}

\title{Group 3 Project Part 1}
\author{
    Kayla Wankowski (kmw32@students.uwf.edu)\\
    Muddasar Ali (ma229@students.uwf.edu)\\
    Raffaele Galliera (rg101@students.uwf.edu)\\
    Tobias Jacob (tj75@students.uwf.edu)
}

\maketitle

\section{Preprocessing steps}
The dataset measures 967 MB in size and contains multiple tables.
We are most interested in the PlayerAttributes, TeamAttributes, and Game table with 37, 7, and 149 columns, respectively.
Our goal is to build an interesting data set that allows us to answer multiple questions.
We created a preprocessing script that takes the full database and reduces it, both in terms of attributes we found interesting and rows.
The output of the script shows the data columns we decided to keep.
We utilized stepwise forward selection to select suitable columns.
The columns and rows we kept are shown in the script output.
For \texttt{Game}, we wanted to keep a way to identify the games, the involved teams, when it happened and how many points each team scored, as well as how many shots have been taken during the game and how many were scored. The attributes related to the score of each team are granular and divided by Field Goals, 3-Pointers and Free throws, both for attempted and made shots.
For \texttt{PlayerAttributes}, we decided to keep the player information, the points scored, the active season and some biometric data like height, weight.
For \texttt{TeamAttributes}, we decided to keep the team information and the founding year.
These attributes allow us to answer questions like: How old are Players typically? At which age/height do the score best? How the presence of a player influences its team in terms of Field Goals Made and Field Goals attempted?

Some of the rows were containing missing values, mostly because old games do not have some of the attributes recorded. As our dataset is quiet abundant, we decided to drop the rows with missing values on the attributes we found interesting and for which we filtered our columns, in order to avoid an injection of bias to our dataset. All the details about the filtering operations are present in the script output below. 

A we mentioned in the last email sent to Professor Bagui, we have decided to not drop the number of rows under 1500. One of the most interesting properties of our dataset is that is indeed complete in the historical records of the games, giving it a drastic cut would, instead of simplifying things, complicate it for us. The final tables contain 41456, 4384, 30 rows respectively for Game, PlayerAttributes and TeamAttributes.

The script reduces the size of the database to 7MB.

\subsubsection*{Output of the script}

\begin{lstlisting}
---Reducing Game ---

Columns before:  GAME_ID, SEASON_ID, TEAM_ID_HOME, TEAM_ABBREVIATION_HOME, TEAM_NAME_HOME, GAME_DATE, MATCHUP_HOME, WL_HOME, MIN_HOME, FGM_HOME, FGA_HOME, FG_PCT_HOME, FG3M_HOME, FG3A_HOME, FG3_PCT_HOME, FTM_HOME, FTA_HOME, FT_PCT_HOME, OREB_HOME, DREB_HOME, REB_HOME, AST_HOME, STL_HOME, BLK_HOME, TOV_HOME, PF_HOME, PTS_HOME, PLUS_MINUS_HOME, VIDEO_AVAILABLE_HOME, TEAM_ID_AWAY, TEAM_ABBREVIATION_AWAY, TEAM_NAME_AWAY, MATCHUP_AWAY, WL_AWAY, MIN_AWAY, FGM_AWAY, FGA_AWAY, FG_PCT_AWAY, FG3M_AWAY, FG3A_AWAY, FG3_PCT_AWAY, FTM_AWAY, FTA_AWAY, FT_PCT_AWAY, OREB_AWAY, DREB_AWAY, REB_AWAY, AST_AWAY, STL_AWAY, BLK_AWAY, TOV_AWAY, PF_AWAY, PTS_AWAY, PLUS_MINUS_AWAY, VIDEO_AVAILABLE_AWAY, GAME_DATE_EST, GAME_SEQUENCE, GAME_STATUS_ID, GAME_STATUS_TEXT, GAMECODE, HOME_TEAM_ID, VISITOR_TEAM_ID, SEASON, LIVE_PERIOD, LIVE_PC_TIME, NATL_TV_BROADCASTER_ABBREVIATION, LIVE_PERIOD_TIME_BCAST, WH_STATUS, TEAM_CITY_HOME, PTS_PAINT_HOME, PTS_2ND_CHANCE_HOME, PTS_FB_HOME, LARGEST_LEAD_HOME, LEAD_CHANGES_HOME, TIMES_TIED_HOME, TEAM_TURNOVERS_HOME, TOTAL_TURNOVERS_HOME, TEAM_REBOUNDS_HOME, PTS_OFF_TO_HOME, TEAM_CITY_AWAY, PTS_PAINT_AWAY, PTS_2ND_CHANCE_AWAY, PTS_FB_AWAY, LARGEST_LEAD_AWAY, LEAD_CHANGES_AWAY, TIMES_TIED_AWAY, TEAM_TURNOVERS_AWAY, TOTAL_TURNOVERS_AWAY, TEAM_REBOUNDS_AWAY, PTS_OFF_TO_AWAY, LEAGUE_ID, GAME_DATE_DAY, ATTENDANCE, GAME_TIME, TEAM_CITY_NAME_HOME, TEAM_NICKNAME_HOME, TEAM_WINS_LOSSES_HOME, PTS_QTR1_HOME, PTS_QTR2_HOME, PTS_QTR3_HOME, PTS_QTR4_HOME, PTS_OT1_HOME, PTS_OT2_HOME, PTS_OT3_HOME, PTS_OT4_HOME, PTS_OT5_HOME, PTS_OT6_HOME, PTS_OT7_HOME, PTS_OT8_HOME, PTS_OT9_HOME, PTS_OT10_HOME, PTS_HOME_y, TEAM_CITY_NAME_AWAY, TEAM_NICKNAME_AWAY, TEAM_WINS_LOSSES_AWAY, PTS_QTR1_AWAY, PTS_QTR2_AWAY, PTS_QTR3_AWAY, PTS_QTR4_AWAY, PTS_OT1_AWAY, PTS_OT2_AWAY, PTS_OT3_AWAY, PTS_OT4_AWAY, PTS_OT5_AWAY, PTS_OT6_AWAY, PTS_OT7_AWAY, PTS_OT8_AWAY, PTS_OT9_AWAY, PTS_OT10_AWAY, LAST_GAME_ID, LAST_GAME_DATE_EST, LAST_GAME_HOME_TEAM_ID, LAST_GAME_HOME_TEAM_CITY, LAST_GAME_HOME_TEAM_NAME, LAST_GAME_HOME_TEAM_ABBREVIATION, LAST_GAME_HOME_TEAM_POINTS, LAST_GAME_VISITOR_TEAM_ID, LAST_GAME_VISITOR_TEAM_CITY, LAST_GAME_VISITOR_TEAM_NAME, LAST_GAME_VISITOR_TEAM_CITY1, LAST_GAME_VISITOR_TEAM_POINTS, HOME_TEAM_WINS, HOME_TEAM_LOSSES, SERIES_LEADER, VIDEO_AVAILABLE_FLAG, PT_AVAILABLE, PT_XYZ_AVAILABLE, HUSTLE_STATUS, HISTORICAL_STATUS 

Columns after reduction:  TEAM_ID_HOME, TEAM_NAME_HOME, TEAM_ID_AWAY, TEAM_NAME_AWAY, GAME_DATE, SEASON, HOME_TEAM_WINS, HOME_TEAM_LOSSES, PTS_HOME, PTS_AWAY, FGM_HOME, FGA_HOME, FG3M_HOME, FG3A_HOME, FTM_HOME, FTA_HOME, FGM_AWAY, FGA_AWAY, FG3M_AWAY, FG3A_AWAY, FTM_AWAY, FTA_AWAY 

Total number of rows:  62448 

Null values found:  TEAM_ID_HOME            0
TEAM_NAME_HOME          0
TEAM_ID_AWAY            0
TEAM_NAME_AWAY          0
GAME_DATE               0
SEASON                  0
HOME_TEAM_WINS         69
HOME_TEAM_LOSSES       69
PTS_HOME                0
PTS_AWAY                0
FGM_HOME               18
FGA_HOME            17862
FG3M_HOME           15657
FG3A_HOME           20921
FTM_HOME               21
FTA_HOME             3036
FGM_AWAY               15
FGA_AWAY            17861
FG3M_AWAY           15657
FG3A_AWAY           20921
FTM_AWAY               15
FTA_AWAY             3033
dtype: int64 

Total number of rows after dropna:  41456 

Null values found after dropna (testing):  TEAM_ID_HOME        0
TEAM_NAME_HOME      0
TEAM_ID_AWAY        0
TEAM_NAME_AWAY      0
GAME_DATE           0
SEASON              0
HOME_TEAM_WINS      0
HOME_TEAM_LOSSES    0
PTS_HOME            0
PTS_AWAY            0
FGM_HOME            0
FGA_HOME            0
FG3M_HOME           0
FG3A_HOME           0
FTM_HOME            0
FTA_HOME            0
FGM_AWAY            0
FGA_AWAY            0
FG3M_AWAY           0
FG3A_AWAY           0
FTM_AWAY            0
FTA_AWAY            0
dtype: int64 

---Reducing Player_Attributes ---

Columns before:  ID, FIRST_NAME, LAST_NAME, DISPLAY_FIRST_LAST, DISPLAY_LAST_COMMA_FIRST, DISPLAY_FI_LAST, PLAYER_SLUG, BIRTHDATE, SCHOOL, COUNTRY, LAST_AFFILIATION, HEIGHT, WEIGHT, SEASON_EXP, JERSEY, POSITION, ROSTERSTATUS, GAMES_PLAYED_CURRENT_SEASON_FLAG, TEAM_ID, TEAM_NAME, TEAM_ABBREVIATION, TEAM_CODE, TEAM_CITY, PLAYERCODE, FROM_YEAR, TO_YEAR, DLEAGUE_FLAG, NBA_FLAG, GAMES_PLAYED_FLAG, DRAFT_YEAR, DRAFT_ROUND, DRAFT_NUMBER, PTS, AST, REB, ALL_STAR_APPEARANCES, PIE 

Columns after reduction:  ID, FIRST_NAME, LAST_NAME, BIRTHDATE, HEIGHT, WEIGHT, PTS, TEAM_ID, TEAM_NAME, FROM_YEAR, TO_YEAR 

Total number of rows:  4500 

Null values found:  ID              0
FIRST_NAME      0
LAST_NAME       0
BIRTHDATE       0
HEIGHT         97
WEIGHT        101
PTS            15
TEAM_ID         0
TEAM_NAME       0
FROM_YEAR       0
TO_YEAR         0
dtype: int64 

Total number of rows after dropna:  4384 

Null values found after dropna (testing):  ID            0
FIRST_NAME    0
LAST_NAME     0
BIRTHDATE     0
HEIGHT        0
WEIGHT        0
PTS           0
TEAM_ID       0
TEAM_NAME     0
FROM_YEAR     0
TO_YEAR       0
dtype: int64 

---Reducing Team_Attributes ---

Columns before:  ID, ABBREVIATION, NICKNAME, YEARFOUNDED, CITY, ARENA, ARENACAPACITY, OWNER, GENERALMANAGER, HEADCOACH, DLEAGUEAFFILIATION, FACEBOOK_WEBSITE_LINK, INSTAGRAM_WEBSITE_LINK, TWITTER_WEBSITE_LINK 

Columns after reduction:  ID, YEARFOUNDED, CITY, NICKNAME 

Total number of rows:  30 

Null values found:  ID             0
YEARFOUNDED    0
CITY           0
NICKNAME       0
dtype: int64 

Total number of rows after dropna:  30 

Null values found after dropna (testing):  ID             0
YEARFOUNDED    0
CITY           0
NICKNAME       0
dtype: int64 
\end{lstlisting}

\subsubsection*{The script}

\begin{minted}[breaklines]{python}
# Import frameworks
import numpy as np
import matplotlib.pyplot as plt
import pandas as pd
import sqlite3

# Connect to databases
con = sqlite3.connect('basketball.sqlite') # Read in data
conOut = sqlite3.connect('reducedDataset.sqlite') # Write data

# Convert GameDatabase
def reduceData(tableName: str, columnsToKeep: str):
    df = pd.read_sql(f'SELECT * FROM {tableName} LIMIT 10', con) # Make a small request with all columns
    print("---Reducing", tableName, "---\n")
    print("Columns before: ", ", ".join([f"{colName}" for (colName, val) in zip(df.columns, df.iloc[1])]), "\n")
    print("Columns after reduction: ", columnsToKeep, "\n")
    df = pd.read_sql(f'SELECT {columnsToKeep} FROM {tableName}', con) # Request only columnsToKeep
    print("Total number of rows: ", df.shape[0], "\n") # Print the number of rows 
    print("Null values found: ", df.isnull().sum(), "\n") # Count null values found by attribute
    df = df.dropna() # Drop rows with missing values
    print("Total number of rows after dropna: ", df.shape[0], "\n")
    print("Null values found after dropna (testing): ", df.isnull().sum(), "\n")
    df.to_sql(tableName, conOut, if_exists = "replace")
    
# convert Game
columnsToKeep = "TEAM_ID_HOME, TEAM_NAME_HOME, TEAM_ID_AWAY, TEAM_NAME_AWAY, GAME_DATE, SEASON, HOME_TEAM_WINS, HOME_TEAM_LOSSES, PTS_HOME, PTS_AWAY, FGM_HOME, FGA_HOME, FG3M_HOME, FG3A_HOME, FTM_HOME, FTA_HOME, FGM_AWAY, FGA_AWAY, FG3M_AWAY, FG3A_AWAY, FTM_AWAY, FTA_AWAY"
reduceData("Game", columnsToKeep)

# convert Player_Attributes
columnsToKeep = "ID, FIRST_NAME, LAST_NAME, BIRTHDATE, HEIGHT, WEIGHT, PTS, TEAM_ID, TEAM_NAME, FROM_YEAR, TO_YEAR"
reduceData("Player_Attributes", columnsToKeep)

# convert Team_Attributes
columnsToKeep = "ID, YEARFOUNDED, CITY, NICKNAME"
reduceData("Team_Attributes", columnsToKeep)

\end{minted}


\end{document}
